THERMOECONOMIC ANALYSIS OF BIODIESEL PRODUCTION FROM USED COOKING OILS

### 2015-05-04 paper.tex
Version including comments of reviewers

% Add local macro commands
\newcommand{\kfame}{kg{\footnotesize ~FAME}}
\newcommand{\mopcr}[1]{\ensuremath{\left\langle {\textbf{#1}^\ast } \right|}}
\newcommand{\tmopcr}[1]{\ensuremath{\prescript{\mathrm{T}}{}{\left\langle {\textbf{#1}^\ast } \right|}}}
\newcommand{\mbr}[1]{\ensuremath{\langle \textbf{#1} \rangle}}
\newcommand{\tmbr}[1]{\ensuremath{\prescript{\mathrm{T}}{}{\langle \textbf{#1} \rangle}}}
\newcommand{\vm}[1]{\ensuremath{\mathbf{#1}}}
\newcommand{\tvm}[1]{\ensuremath{\prescript{\mathrm{T}}{}{\mathbf{#1}}}}
\newcommand{\vms}[3][\phantom{\ast}]{\ensuremath{\mathbf{#2}_{#3}^\mathrm{#1}}}
\newcommand{\ems}[3][\phantom{\ast}]{\ensuremath{{#2}_{#3}^\mathrm{#1}}}
\newcommand{\degreeC}[1]{#1~\textcelsius}
\newcolumntype{W}{@{}>{\raggedright\arraybackslash}X@{}}
\newcolumntype{Y}{@{}>{\raggedleft\arraybackslash}X@{}}
\newcolumntype{Z}{@{}>{\centering\arraybackslash}X@{}}

@@

With the current legislation in place in the European Union the use of UCO for biodiesel production has gained importance. According to article 21 of the Renewable Energy Directive \cite{Directive2009/28/EC}, the contribution made by biofuels to the 2020 targets produced from wastes, residues, non-food cellulosic material, and ligno-cellulosic material are considered to be twice that made by other biofuels. Biofuels produced from these resources are called advanced or second generation biofuels. Biodiesel from UCO is one of them.

This regulation has had unexpected counter effects \cite{Bailey2013,ePURE2013,EBTP2011} international imports of UCO have started to take place, the price of UCO has scaled up to even exceed the price of edible vegetable oils, such as palm oil, and finally it has opened the door to fraud. As there is no real definition of used cooking oil and no possible way nowadays to determine when a vegetable oil is used or not, some actors are claiming to have UCO when in reality the vegetable oil has only been used once and they have used high quantities of oil to fry a small portion of food. In light of the revision of the Renewable Energy Directive, there are voices suggesting eliminating UCO from the list of advanced biofuels benefiting from double counting \cite{ePURE2013}.

@@

Being this the situation, some projects have estimated the potential biodiesel production in the EU. The IEE project RecOil states that biodiesel produced from UCO could replace 1.5\% of the EU-27 diesel consumption \cite{RecOil2013}. The final report of the IEE project BioDieNet \cite{BioDieNet2009} states that the total amount of recoverable UCO available in the EU-27 in 2009 was about 3.5 million tonnes (3.95 million litres). As the EU-27 diesel consumption was 214 million litres in 2007, around 1.8\% of the EU-27 diesel consumption could have potentially been replaced with biodiesel produced from UCO.

@@

The life cycle starts by the collection from households, HORECA and industry and is followed by the recycling where the oil is filtered and decanted in order to separate solid particles and water. Once the oil is refined, it is sent to the transesterification plant.

@@

The collection can be organised in different ways. In the HORECA sector and food industry, normally the collection is done on-site by the authorised agent in medium-to-big containers that are transported via ban or truck. In the case of households there are different possibilities: door-to door collection or distributed collection in points throughout an area with well identified containers. These containers can be adapted to receive UCO in recipients (such as used bricks or bottles, or reusable bottles designed for the purpose) or have a hole where the oil can be poured through. According to results of IEE projects OILECO and RecOiL the best practices, showing higher acceptance, are those having collection points with containers adapted to leave recipients \cite{OILECO2013,RecOil2013}. This system can also be the most sustainable if the transport of the oil to the collection point by the household consumer is not accounted in the life cycle assessment.

@@

As far as the transesterification is concerned, the UCO normally follows a different process compared to the case of crude vegetable oils, which affects the consumption of materials and energy sources. The main differences are:
\begin{itemize}
\item The pre-treatment is not so elaborated because the oil comes already ``clean'' from the recycling unit. Normally, only a cleaning is done via centrifugation and drying to eliminate possible small particles and excess water.
\item Afterwards, the oil is esterified with an acid catalyser to, as commented above, transform into esters the high quantity of free fatty acids that are normally present in UCO, this esterification is followed by a basic catalysed transesterification. Some plants have the esterification and transesterification tanks in parallel. They treat in the esterification plant those oils having high FFA (more than 4\%) and in the transesterification plant those oils with less FFA \cite{CIEMAT2005}.
\item The ester obtained follows a cleaning and drying process. Other plants have opted to distillate it to ensure the fulfilment of the quality standards, EN 14214 \cite{CENEN14214} in the EU.
\item The glycerol does not follow an after-treatment process. The crude glycerol is either treated as a residue or sold as a low added-value co-product to soap unfolders or to be used as fuel in cement kilns \cite{FontdeMoraThesis2013}. Some plants have after-treatment processes similar to the ones using vegetable oils to obtain a high quality glycerine product; these after-treatment processes normally consist in adding an acid, e.g. HCl, to split the soaps into FFA and salts which can then be separated by decantation, and distilling the remaining flow up to \degreeC{130} to evaporate the remaining methanol \cite{FontdeMoraThesis2013}.
\end{itemize}

@@

\begin{figure}[htbp]
\centering
\includegraphics[width=0.6\textwidth]{Figure1}
\caption[Flow diagram of a biodiesel plant producing with Used Cooking Oil]{Flow diagram of a biodiesel plant producing with Used Cooking Oil. Source: CIEMAT, 2005}
\label{fig1}
\end{figure}

@@

\begin{table}[htbp]
\centering \small
\caption[Consumption data for processes considered in the life cycle analysis for UCO]{Consumption data for processes considered in the life cycle analysis for UCO. Source: CIEMAT, 2005}
\begin{tabularx}{0.8\textwidth}{lXXXX}
\toprule
\multicolumn{2}{l}{\bf Process} & {\bf I/O}   & {\bf Value} & {\bf Units} \\
\midrule
\multicolumn{5}{l}{\bf Collection} \\
\cmidrule{1-5}
& Diesel & Input & 0.48316 & MJ/kg UCO \\
\cmidrule{1-5}
\multicolumn{5}{l}{\bf Recycling} \\
\cmidrule{1-5}
& Diesel & Input & 0.036 & MJ/kg oil \\
& Electricity & Input & 0.35  & MJ/kg oil \\
\cmidrule{1-5}
\multicolumn{5}{l}{\bf Transesterification} \\
\cmidrule{1-5}
& Recycled oil& Input& 1.03& kg/\kfame \\
& Methanol& Input& 0.155& kg/\kfame \\
& KOH& Input& 0.02770& kg/\kfame \\
& \ce{H2SO4}& Input& 0.017& kg/\kfame \\
& Diesel& Input& 0.455& kWh/\kfame \\
& Electricity & Input& 0.137& kWh/\kfame \\
& FAME & Output & 1.00& kg/\kfame \\
& Glycerol crude& Output& 0.20& kg/\kfame \\
\bottomrule
\end{tabularx}
\label{table1}
\end{table}

@@

The methodology applied in this paper has been introduced in previous works \cite{FontdeMora2012,FontdeMora2013,Torres2012b}. Thermoeconomic Input--Output Analysis, formerly Symbolic Exergoeconomics \cite{Torres2006}, provides general relationships between the production demand and the resources cost with the efficiency and irreversibilities of each individual process in an energy system. The distinguishing element in thermoeconomics from conventional energy and exergy analysis is purpose. Matter and energy flows entering and exiting a given system are classified into fuel and product. Fuel (F) refers to the resources that the component uses to achieve its purpose, and product (P) corresponds to the flows related to that purpose.

The methodology is closely related to the Input-Output analysis \cite{Miller2009}. The mathematical principles are very similar, but the input-output table is transformed into a Fuel-Product model, in which the Second Law is used in the analysis of the processes. The exergy cost of the system are calculated verifying the properties of the exergy cost theory \cite{Lozano1993}. The exergy cost analysis refers to consumed resources into the boundary limits of the system. The costs of the external resources are equal to its exergy values.

The productive structure represented by the Fuel--Product table, see \cref{table2}, which describes how the production processes are related.

@@

The exergy cost of the products of the system processes, $C_{P,i}$ is the sum of the cost of the resources consumed by the process: system internals, i.e. products of other processes $C_{ji}$, and external resources $C_{0,i}$.

@@

\noindent and the costs of the products obtained in a process are proportional to its exergy:

@@

\begin{equation}
C_{P,i}-\sum_{j=1}^{n}{y_{ji}C_{P,j}}=C_{0i} \qquad i=1,..n
\end{equation}

\noindent which could be written in matrix form as:

\begin{equation}
 \vm{C}_P =\tmopcr{P} \, \vm{C}_e \qquad \text{where} \qquad \mopcr{P} \equiv \left(\vm{U}-\mbr{FP}\right)^{-1}
 \label{eq4}
\end{equation}

@@

The production cost can be decomposed \cite{Torres2012b}, considering the different types of resources required. The resources cost vector can be separated into two terms:

@@

The first term represents the cumulative exergy of the renewable resources and the latter the non-renewable of fossil resources. Therefore, the exergoecologic cost can be broken down into the renewable and fossil parts: $\vms{C}{P}=\vms[rs]{C}{P}+\vms[nrs]{C}{P}$, where:

@@

This distinction distinction depending on the origin of the resources allow us calculate the sustainability indicator. Font de Mora et al. \cite{FontdeMora2012} defined the concept of ExROI and renewability. ExROI measures the amount of product obtained with one unit of non-renewable resources:

@@

\begin{equation}
\label{rho}
\rho=\frac{\ems[rs]{C}{P}}{\ems{C}{P}} \quad 0 \leq \rho \leq 1
\end{equation}

the portion of renewable resources per total resources used, or the weight of the renewable exergy cost with respect to the total cost.

\Cref{table3} shows the exergy of flows an cumulative exergy of external resources of the UCO biodiesel pathway. From this information the F--P table of the UCO pathway model is build, see \cref{table4}.

\begin{sidewaystable}[htbp]
  \centering \small
  \caption{Exergy and cost of flows of the UCO biodiesel pathway}
    \begin{tabularx}{0.9\textwidth}{lcccYYYYY}
    \toprule
          &      &   &    &       & \multicolumn{2}{c}{\bf Exergy} & \multicolumn{2}{c}{\bf Cost} \\
    \cmidrule(r){6-7} \cmidrule(r){8-9}
    {\bf Material} & {\bf Process} & {\bf I/O} & {\bf Unit} (u) & u/\kfame & MJ/u  & MJ/\kfame & MJ/u  & MJ/\kfame \\
    \midrule
    Diesel & 1     & Input & MJ    & 0.6459 & 1.0000 & 0.6459 & 1.1773 & 0.7604 \\
    UCO   & 1     & Output & kg    & 1.3368 & 39.5994 & 52.9364 &  -     & - \\
    Diesel & 2     & Input & MJ    & 0.0371 & 1.0000 & 0.0371 & 1.1773 & 0.0437 \\
    Electricity & 2     & Input & MJ    & 0.4679 & 1.0000 & 0.4679 & 2.8690 & 1.3423 \\
    UCO recycled & 2     & Output & kg    & 1.0300 & 39.5994 & 40.7873 &  -     & - \\
    Methanol & 3     & Input & kg    & 0.1550 & 23.1540 & 3.5889 & 30.3074 & 4.6977 \\
    KOH   & 3     & Input & kg    & 0.0277 & 0.8229 & 0.0228 & 13.3852 & 0.3708 \\
    \ce{H2SO4} & 3     & Input & kg    & 0.0170 & 1.6417 & 0.0279 & 2.6295 & 0.0447 \\
    Diesel & 3     & Input & MJ    & 1.5651 & 1.0000 & 1.5651 & 1.1773 & 1.8426 \\
    Electricity & 3     & Input & MJ    & 0.4713 & 1.0000 & 0.4713 & 2.8690 & 1.3522 \\
    Crude FAME & 3     & Output & kg    & 1.0100 & 39.8548 & 40.2533 &  -     & - \\
    Crude Glycerol & 3     & Output & kg    & 0.2000 & 23.6779 & 4.7356 &    -   & - \\
    HCl   & 4     & Input & kg    & 0.0406 & 1.3409 & 0.0544 & 5.2484 & 0.2131 \\
    Diesel & 4     & Input & MJ    & 0.0749 & 1     & 0.0749 & 1.1773 & 0.0882 \\
    Electricity & 4     & Input & MJ    & 0.0187 & 1.0000 & 0.0187 & 2.869 & 0.0535 \\
    FAME  & 4     & Output & kg    & 1     & 39.8376 & 39.8376 & -      & - \\
    \bottomrule
    \end{tabularx}%
  \label{table3}%
\end{sidewaystable}%

\begin{figure}[htpb]
\centering
\includegraphics[width=0.95\textwidth]{Figure2}
\caption{Cost diagram of UCO biodiesel production}
\label{fig2}
\end{figure}

The results of the cost analysis are shown is \cref{table5} and represented graphically in \cref{fig2}. From these values it is possible to compute the ExROI and renewability factor, whose are 4.10 and 0.83, respectively. This means that from each unit of non-renewable sources, i.e. fossil fuels, it is possible to obtain 4.10 units of energy and that the percentage of renewable energy contained in the fuel is about 83\%. These values can be increased if the conventional methanol of fossil origin is substituted by a bioproduct, either bioethanol or methanol from biomass.

\begin{table}[htbp]
\centering 
\caption[F--P Tables UCO biodiesel]{F--P Tables UCO biodiesel (MJ/kg FAME)}
\begin{tabularx}{0.9\textwidth}{ZZZZZZZ}
\toprule
 & $F_0$  & $F_1$ & $F_2$ & $F_3$ & $F_4$ & Total \\
\midrule
% Table generated by Excel2LaTeX from sheet 'Cost UCO'
\ems[rs]{C}{0} &       & 52.94 & 0.00  & 0.00  & 0.00  & 52.94 \\
\ems[nrs]{C}{0} &       & 0.76  & 1.39  & 8.31  & 0.35  & 10.81 \\
\cmidrule{1-7}
$P_1$ & 0.00  & 0.00  & 52.94 & 0.00  & 0.00  & 52.94 \\
$P_2$ & 0.00  & 0.00  & 0.00  & 40.79 & 0.00  & 40.79 \\
$P_3$ & 4.74  & 0.00  & 0.00  & 0.00  & 40.25 & 44.99 \\
$P_4$ & 39.84 & 0.00  & 0.00  & 0.00  & 0.00  & 39.84 \\
\midrule
Total & 44.58 & 53.70 & 54.32 & 49.10 & 40.61 &  \\
\bottomrule
\end{tabularx}%
\label{table4}%
\end{table}%

\begin{table}[htbp]
  \centering
  \caption[Cost analysis for UCO biodisel]{Cost analysis for UCO biodisel (MJ/kg FAME)}
    \begin{tabularx}{0.8\textwidth}{clZZZZ}
    \toprule
        &                       & $C_P$  & $\ems[rs]{C}{P}$  & $\ems[nrs]{C}{P}$ & $\rho$ \\
    Nr  & Process               & MJ/kg  & MJ/Kg & MJ/kg & \% \\
    \midrule
    1   & Collection            & 53.697 & 52.936 &  0.760 & 98.58 \\
    2   & Refining              & 55.083 & 52.936 &  2.146 & 96.10 \\
    3   & Esterification        & 63.391 & 52.936 & 10.454 & 83.32 \\
    4   & FAME after treatment  & 57.073 & 47.364 &  9.836 & 82.80 \\
    \bottomrule
    \end{tabularx}%
  \label{table5}%
\end{table}%

The values obtained for the UCO biodiesel can be compared to the ones obtained for biodiesels from vegetable oils in \cref{fig3}.  The values for vegetable oils were calculated in \cite{FontdeMoraThesis2013}. As it can be seen, biodiesel production from UCO has the highest ExROI value and the second highest renewability factor, only exceeded by palm oil based biodiesel. The high renewability of palm oil is due to the use of biomass resources to cover part of the energy needs in the cycle. In addition, UCO biodiesel has the lowest non-renewable exergoecologic cost, which means that overall, it consumes less non-renewable resources, i.e. mainly fossil fuels, than its counterparts.

>>>>

% Add local macro commands
\newcommand{\kfame}{kg{\footnotesize ~FAME}}
\newcommand{\mopcr}[1]{\ensuremath{\left\langle {\textbf{#1}^\ast } \right|}}
\newcommand{\tmopcr}[1]{\ensuremath{\prescript{\mathrm{T}}{}{\left\langle {\textbf{#1}^\ast } \right|}}}
\newcommand{\mbr}[1]{\ensuremath{\langle \textbf{#1} \rangle}}
\newcommand{\tmbr}[1]{\ensuremath{\prescript{\mathrm{T}}{}{\langle \textbf{#1} \rangle}}}
\newcommand{\vm}[1]{\ensuremath{\mathbf{#1}}}
\newcommand{\tvm}[1]{\ensuremath{\prescript{\mathrm{T}}{}{\mathbf{#1}}}}
\newcommand{\vms}[3][\phantom{\ast}]{\ensuremath{\mathbf{#2}_{#3}^\mathrm{#1}}}
\newcommand{\ems}[3][\phantom{\ast}]{\ensuremath{{#2}_{#3}^\mathrm{#1}}}
\newcommand{\elm}{\ensuremath{\left[\vm{A}_P^e\right]}}

\newcommand{\degreeC}[1]{#1~\textcelsius}
\newcolumntype{W}{@{}>{\raggedright\arraybackslash}X@{}}
\newcolumntype{Y}{@{}>{\raggedleft\arraybackslash}X@{}}
\newcolumntype{Z}{@{}>{\centering\arraybackslash}X@{}}

@@

With the current legislation in place in the European Union the use of UCO for biodiesel production has gained importance. According to article 21 of the Renewable Energy Directive \cite{Directive2009/28/EC}, the contribution made by biofuels to the 2020 targets produced from wastes, residues, non-food cellulosic material, and ligno-cellulosic material are considered to be twice that made by other biofuels. Biofuels produced from these resources are called advanced or second generation biofuels. Biodiesel from UCO is one of them. The importance of UCO derived biodiesel will increase even more when a limit to the production of crop based biofuels will be introduced in the European legislation as recently agreed by the European Parliament and the Member States.

This regulation has had unexpected counter effects \cite{Bailey2013,ePURE2013,EBTP2011} as international imports of UCO have started to take place, the price of UCO has scaled up to even exceed the price of edible vegetable oils, such as palm oil, and finally it has opened the door to fraud. As there is no real definition of used cooking oil and no possible way nowadays to determine when a vegetable oil is used or not, some actors are claiming to have UCO when in reality the vegetable oil has only been used once and they have used high quantities of oil to fry a small portion of food. In light of the revision of the Renewable Energy Directive, there are voices suggesting eliminating UCO from the list of advanced biofuels benefiting from double counting \cite{ePURE2013}.

@@

Being this the situation, some projects have estimated the potential biodiesel production in the EU. The IEE project RecOil states that biodiesel produced from UCO could replace 1.5\% of the EU-27 diesel consumption \cite{RecOil2013}. The final report of the IEE project BioDieNet \cite{BioDieNet2009} states that the total amount of recoverable UCO available in the EU-27 in 2009 was about 3.5 million tonnes (3,950 million litres). As the EU-27 diesel consumption was 214 billion litres in 2007, around 1.8\% of the EU-27 diesel consumption could have potentially been replaced with biodiesel produced from UCO.

@@

The life cycle, see \cref{fig1}, starts by the collection from households, HORECA and industry and is followed by the recycling where the oil is filtered and decanted in order to separate solid particles and water. Once the oil is refined, it is sent to the transesterification plant.

@@

The collection can be organised in different ways. In the HORECA sector and food industry, normally the collection is done on-site by the authorised agent in medium-to-big containers that are transported via van or truck. In the case of households there are different possibilities: door-to door collection or distributed collection in points throughout an area with well identified containers. These containers can be adapted to receive UCO in recipients (such as used bricks or bottles, or reusable bottles designed for the purpose) or have a hole where the oil can be poured through. According to results of IEE projects OILECO and RecOiL the best practices, showing higher acceptance, are those having collection points with containers adapted to leave recipients \cite{OILECO2013,RecOil2013}. This system can also be the most sustainable if the transport of the oil to the collection point by the household consumer is not accounted in the life cycle assessment.

@@

As far as the transesterification is concerned, the UCO normally follows a different process compared to the case of crude vegetable oils, which affects the consumption of materials and energy sources. The main differences are:
\begin{itemize}
\item The pre-treatment is not so elaborated because the oil comes already ``clean'' from the recycling unit. Normally, only a cleaning is done via centrifugation and drying to eliminate possible small particles and excess water.
\item Afterwards, the oil is esterified with an acid catalyser to transform into esters the high quantity of free fatty acids that are normally present in UCO, this esterification is followed by a basic catalysed transesterification. Some plants have the esterification and transesterification tanks in parallel. They treat in the esterification plant those oils having high FFA (more than 4\%) and in the transesterification plant those oils with less FFA \cite{CIEMAT2005}.
\item The ester obtained follows a cleaning and drying process. Other plants have opted to distillate it to ensure the fulfilment of the quality standards, EN 14214 \cite{CENEN14214} in the EU.
\item The glycerol does not follow an after-treatment process. The crude glycerol is either treated as a residue or sold as a low added-value co-product to soap unfolders or to be used as fuel in cement kilns \cite{FontdeMoraThesis2013}. Some plants have after-treatment processes similar to the ones using vegetable oils to obtain a high quality glycerine product; these after-treatment processes normally consist in adding an acid, e.g. HCl, to split the soaps into FFA and salts which can then be separated by decantation, and distilling the remaining flow up to \degreeC{130} to evaporate the remaining methanol \cite{FontdeMoraThesis2013}.
\end{itemize}

@@

\begin{figure}[htbp]
\centering
\includegraphics[width=0.5\textwidth]{Figure1}
\caption[Flow diagram of a biodiesel plant producing with Used Cooking Oil]{Flow diagram of a biodiesel plant producing with Used Cooking Oil. Source: CIEMAT \cite{CIEMAT2005}}
\label{fig1}
\end{figure}

@@

\begin{table}[htbp]
\centering \small
\caption[Consumption data for processes considered in the life cycle analysis for UCO]{Consumption data for processes considered in the life cycle analysis for UCO. Source: CIEMAT \cite{CIEMAT2005}}
\begin{tabularx}{0.8\textwidth}{lXXXX}
\toprule
\multicolumn{2}{l}{\bf Process} & {\bf I/O}   & {\bf Value} & {\bf Units} \\
\midrule
\multicolumn{5}{l}{\bf Collection} \\
\cmidrule{1-5}
& Diesel & Input & 0.48316 & MJ/kg UCO \\
\cmidrule{1-5}
\multicolumn{5}{l}{\bf Recycling} \\
\cmidrule{1-5}
& Diesel & Input & 0.036 & MJ/kg oil \\
& Electricity & Input & 0.35  & MJ/kg oil \\
\cmidrule{1-5}
\multicolumn{5}{l}{\bf Transesterification} \\
\cmidrule{1-5}
& Recycled oil& Input& 1.03& kg/\kfame \\
& Methanol& Input& 0.155& kg/\kfame \\
& KOH& Input& 0.02770& kg/\kfame \\
& \ce{H2SO4}& Input& 0.017& kg/\kfame \\
& Diesel& Input& 0.455& kWh/\kfame \\
& Electricity & Input& 0.137& kWh/\kfame \\
& FAME & Output & 1.00& kg/\kfame \\
& Glycerol crude& Output& 0.20& kg/\kfame \\
\bottomrule
\end{tabularx}
\label{table1}
\end{table}

@@

The methodology applied in this paper has been introduced in previous works \cite{FontdeMora2012,FontdeMora2013,Torres2012b}. Thermoeconomic Input--Output Analysis, formerly Symbolic Exergoeconomics \cite{Torres2006}, applies the mathematical framework of Input--Output analysis [20] to Exergy Cost Theory \cite{Miller2009}. It provides general relationships between the production demand and the resources cost of an energy system with the efficiency and irreversibilities of each individual process of that system. The distinguishing element in thermoeconomics from conventional energy and exergy analysis is purpose. Matter and energy flows entering and exiting a given system are classified into fuel and product. Fuel (F) refers to the resources that the component uses to achieve its purpose, and product (P) corresponds to the flows related to that purpose \cite{Tsatsaronis1985}.

The system interactions are represented by means of the fuel--product table, see \cref{table2}, which describes how the production processes are related.

@@

Let us consider a system formed by, say $n$ processes, and the environment named as process number 0. Let $P_i$ denote the  exergy of the process $i$ production, which is used in part to meet the intermediate requirement as input resources of other processes and in part to meet the final demand of the system. If $E_{ij}$ denotes the exergy of process $i$ uses as resource for process $j$, and $E_{i0}$ denotes the final demand produced in process $i$, the following expression is formed:
\begin{equation}
\label{eq:F}
P_i=E_{i0}+\sum_{j}^{n}E_{ij}, \quad i=1,\ldots,n
\end{equation}
Note that, in this model, the production of the environment $P_0$ is the total exergy of the external resources. On the other hand, the input resources of each process $i$, say $F_i$, is in
part coming from external resources, say $E_{0i}$ and in part from the
production of other processes, then we have the condition:
\begin{equation}
\label{eq:P}
F_i=E_{0i}+\sum_{j}^{n}E_{ji}, \quad i=1,\ldots,n
\end{equation}
\noindent where $F_0$ is the total system demand.

Second Law states that the difference between fuel and product is a positive 
value and equal to the irreversibility (I), accounting for the losses 
appearing in the process: $F_i - P_i = I_i \geq 0$. Furthermore, the unit 
exergy consumption is defined as the ratio between fuel and product: 
$k_i={F_i}/{P_i} \geq 1$

The exergy cost of the products of a system refers to the exergy of the resources consumed to obtain such products. Therefore the exergy cost of the production of a process, denoted as $C_{P,i}$,  is the sum of the cost of the resources consumed by the process: system internals, i.e. products of other processes $C_{ji}$, and the cumulative exergy of the external resources $C_{0,i}$.

@@

To calculate the production exergy cost, the assumption that the costs of the products obtained in a process are proportional to its exergy is made, see \cite{Lozano1993}:

@@

\begin{equation}
C_{P,i}=C_{0i}+\sum_{j=1}^{n}{y_{ji}C_{P,j}} \qquad i=1,..n
\end{equation}

\noindent which could be written in matrix form as:
\begin{equation}
\vm{C}_P = \vm{C}_e + \tmbr{FP} \vm{C}_P
\end{equation}

@@

Therefore, the production cost could be obtained as:

\begin{equation}
 \vm{C}_P =\tmopcr{P} \, \vm{C}_e \qquad \text{where} \qquad \mopcr{P} \equiv \left(\vm{U}-\mbr{FP}\right)^{-1}
 \label{eq4}
\end{equation}
where \mopcr{P} is called production cost matrix, and its elements $\pi_{ij}^{\ast}$ are the global distribution ratios, whose represent the amount of external resources (measured in exergy) of process $i$ required to obtain one exergy unit of the product of the process $j$.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{sidewaystable}[htbp]
  \centering \small
  \caption{Exergy and cost of flows of the UCO biodiesel pathway}
    \begin{tabularx}{0.9\textwidth}{lcccYYYYY}
    \toprule
          &      &   &    &       & \multicolumn{2}{c}{\bf Exergy} & \multicolumn{2}{c}{\bf Cost} \\
    \cmidrule(r){6-7} \cmidrule(r){8-9}
    {\bf Material} & {\bf Process} & {\bf I/O} & {\bf Unit} (u) & u/\kfame & MJ/u  & MJ/\kfame & MJ/u  & MJ/\kfame \\
    \midrule
    Diesel         & 1  & Input  & MJ    & 0.6459 & 1.0000  & 0.6459  & 1.1773  & 0.7604 \\
    UCO            & 1  & Output & kg    & 1.3368 & 39.5994 & 52.9364 &  -      & - \\
    Diesel         & 2  & Input  & MJ    & 0.0371 & 1.0000  & 0.0371  & 1.1773  & 0.0437 \\
    Electricity    & 2  & Input  & MJ    & 0.4679 & 1.0000  & 0.4679  & 2.8690  & 1.3423 \\
    UCO recycled   & 2  & Output & kg    & 1.0300 & 39.5994 & 40.7873 &  -      & - \\
    Methanol       & 3  & Input  & kg    & 0.1550 & 23.1540 & 3.5889  & 30.3074 & 4.6977 \\
    KOH            & 3  & Input  & kg    & 0.0277 & 0.8229  & 0.0228  & 13.3852 & 0.3708 \\
    \ce{H2SO4}     & 3  & Input  & kg    & 0.0170 & 1.6417  & 0.0279  & 2.6295  & 0.0447 \\
    Diesel         & 3  & Input  & MJ    & 1.5651 & 1.0000  & 1.5651  & 1.1773  & 1.8426 \\
    Electricity    & 3  & Input  & MJ    & 0.4713 & 1.0000  & 0.4713  & 2.8690  & 1.3522 \\
    Crude FAME     & 3  & Output & kg    & 1.0100 & 39.8548 & 40.2533 &  -      & - \\
    Crude Glycerol & 3  & Output & kg    & 0.2000 & 23.6779 & 4.7356  &  -      & - \\
    HCl            & 4  & Input  & kg    & 0.0406 & 1.3409  & 0.0544  & 5.2484  & 0.2131 \\
    Diesel         & 4  & Input  & MJ    & 0.0749 & 1       & 0.0749  & 1.1773  & 0.0882 \\
    Electricity    & 4  & Input  & MJ    & 0.0187 & 1.0000  & 0.0187  & 2.869   & 0.0535 \\
    FAME           & 4  & Output & kg    & 1      & 39.8376 & 39.8376 &  -      & - \\
    \bottomrule
    \end{tabularx}%
  \label{table3}%
\end{sidewaystable}%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\Cref{table3} shows the exergy of flows and cumulative exergy of external resources of the UCO biodiesel pathway. From this information the F--P table of the UCO pathway model is built, see \cref{table4}.

\begin{table}[H]
\centering 
\caption[F--P Tables UCO biodiesel]{F--P Tables UCO biodiesel (MJ/kg FAME)}
\begin{tabularx}{0.9\textwidth}{ZZZZZZZ}
\toprule
 & $F_0$  & $F_1$ & $F_2$ & $F_3$ & $F_4$ & Total \\
\midrule
% Table generated by Excel2LaTeX from sheet 'Cost UCO'
$P_0$ &       & 53.58 & 0.51  & 5.68  & 0.15  & 59.91 \\
\cmidrule{1-7}
$P_1$ & 0.00  & 0.00  & 52.94 & 0.00  & 0.00  & 52.94 \\
$P_2$ & 0.00  & 0.00  & 0.00  & 40.79 & 0.00  & 40.79 \\
$P_3$ & 4.74  & 0.00  & 0.00  & 0.00  & 40.25 & 44.99 \\
$P_4$ & 39.84 & 0.00  & 0.00  & 0.00  & 0.00  & 39.84 \\
\midrule
Total & 44.58 & 53.58 & 53.44 & 46.46 & 40.40 &  \\
\bottomrule
\end{tabularx}%
\label{table4}%
\end{table}%

The local and global distribution matrices, are obtained from \cref{table4} using \cref{eq2,eq4}:
\[
\mbr{FP} = \begin{bmatrix}
0.0000 & 1.0000 & 0.0000 & 0.0000 \bigstrut[t]\\
0.0000 & 0.0000 & 1.0000 & 0.0000 \\
0.0000 & 0.0000 & 0.0000 & 0.8947 \\
0.0000 & 0.0000 & 0.0000 & 0.0000 \bigstrut[b]\\
\end{bmatrix}
\qquad
\mopcr{P} = \begin{bmatrix}
1.0000 & 1.0000 & 1.0000 & 0.8947 \bigstrut[t]\\
0.0000 & 1.0000 & 1.0000 & 0.8947 \\
0.0000 & 0.0000 & 1.0000 & 0.8947 \\
0.0000 & 0.0000 & 0.0000 & 1.0000 \bigstrut[b]\\
\end{bmatrix}
\]

The production exergy cost can be decomposed \cite{Torres2012b}, considering the different types of resources required. The resources cost vector can be separated into two terms:

@@

The first term represents the cumulative exergy of the renewable resources and the latter the non-renewable or fossil resources. Therefore, the exergoecologic cost can be broken down into the renewable and fossil parts: $\vms{C}{P}=\vms[rs]{C}{P}+\vms[nrs]{C}{P}$, where:

@@

This distinction depending on the origin of the resources allows calculating the sustainability indicator. Font de Mora et al. \cite{FontdeMora2012} defined the concept of ExROI and renewability. ExROI measures the amount of product obtained with one unit of non-renewable resources:

@@

\begin{equation}
\label{rho}
\rho=\frac{\ems[rs]{C}{P}}{\ems{C}{P}} \quad 0 \leq \rho \leq 1
\end{equation}
\noindent the portion of renewable resources per total resources used, or the weight of the renewable exergy cost with respect to the total cost.

The results of the cost analysis, obtained by applying \cref{cpnrs},  are shown in \cref{table5}. In this case, the primary UCO stream is considered as a renewable resource and its cost is equal to its exergy. From these values it is possible to compute the ExROI and renewability factor, which are 4.10 and 0.83, respectively. This means that from each unit of non-renewable sources, i.e. fossil fuels, it is possible to obtain 4.10 units of energy and that the percentage of renewable energy contained in the fuel is about 83\%. These values can be increased if the conventional methanol of fossil origin is substituted by a bioproduct, either bioethanol or methanol from biomass.

\begin{table}[htbp]
  \centering
  \caption[Cost analysis for UCO biodisel]{Cost analysis for UCO biodisel (MJ/kg FAME)}
    \begin{tabularx}{0.8\textwidth}{clZZZZZ}
    \toprule
        &                       & $\ems[nrs]{C}{e}$  & $C_P$  & $\ems[rs]{C}{P}$  & $\ems[nrs]{C}{P}$ & $\rho$ \\
    Nr  & Process               & MJ/Kg  & MJ/kg  & MJ/Kg & MJ/kg & \% \\
    \midrule
    1   & Collection            & 0.76   & 53.697 & 52.936 &  0.760 & 98.58 \\
    2   & Refining              & 1.39   & 55.083 & 52.936 &  2.146 & 96.10 \\
    3   & Esterification        & 8.31   & 63.391 & 52.936 & 10.454 & 83.51 \\
    4   & FAME after treatment  & 0.35   & 57.073 & 47.364 &  9.709 & 82.99 \\
    \bottomrule
    \end{tabularx}%
  \label{table5}%
\end{table}%

The values obtained for the UCO biodiesel can be compared to the ones obtained for biodiesels from vegetable oils in \cref{fig3}.  The values for vegetable oils were calculated in \cite{FontdeMoraThesis2013}. As it can be seen, biodiesel production from UCO has the highest ExROI value and the second highest renewability factor, only exceeded by palm oil based biodiesel. The high renewability of palm oil is due to the use of biomass resources to cover part of the energy needs in the cycle. In addition, UCO biodiesel has the lowest non-renewable exergy cost, which means that overall, it consumes less non-renewable resources, i.e. mainly fossil fuels, than its counterparts. The left graphic, in \cref{fig3}, shows also the renewable and non-renewable exergy cost per produced unit of the diferent biodiesels analyzed in the paper. This analysis could be extended to cover also non-edible vegetable oils, for example castor and jatropha oils, and their blends with UCO \cite{Kannan2013}.
